---
title: "eiCompare: Ecological Inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette illustrates how we can use the Ecological Inference (EI) tools 
within `eiCompare` to learn about the voting behavior of different groups of 
voters. 

## The ecological inference problem

Ecological Inference involves estimating the voting behaviors of different 
groups of voters using only data about election results and the number of people
from each group that turned out to vote. Elections in the United States, and 
much of the rest of the world, use secret ballots that ensure confidentiality 
for every voter. Voter confidentiality is important for a functioning democracy, 
but it presents a problem for those who want to learn about voter behavior. We cannot directly 
measure how people vote in an election. Intead, we have to make inferences about how they might have voted 
using election results and information about voter turnout. We often end up with
datasets like the following, which contains aggregate election results and turnout 
information for each election precinct in Gwinnett County, Georgia:

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(eiCompare)

data("gwinnett")
head(gwinnett)
```

This dataset contains the following columns:

- `precinct` - This uniquely identifies each precinct in the county.
- `turnout` - The total number of voters who cast ballots on election day
- `kemp`, `abrams`, `metz` - These count the number of votes cast for each
candidate.
- `white`, `black`, `hispanic`, `other` - These count the number of voters who
self-reported as each race group.

Election results by precinct are generally accessible online from election 
administrators. Measuring the turnout of different racial/ethnic groups, 
however, can present more of a challenge. Luckily, the state of Georgia requires
voters to self-report their race when they register. This means anyone can 
tabulate the number of voters self-reporting with each group in each precinct, 
provided they gain access to a voter file. That's exactly what we've done here.
To learn more about how to estimate voters' race based only on their name and 
address, see the tutorials on [Geocoding voter files](geo.com) and 
[Bayesian Improved Surname Geocoding](BISG.com)

We know the number of votes cast for each candidate and by each racial/ethnic 
group. From this data, we want to learn whether voters from different ethnic
groups prefer different candidates. 

## Preparing the data

First, we need to clean and prepare the data. `eiCompare` contains functions to
streamline data preparation and prevent common errors that emerge during this 
process.

First, let's set up our column vectors. Most `eiCompare` functions for ecological inference require these as parameters.

```{r}
cands <- c("kemp", "abrams", "metz")
races <- c("white", "black", "hispanic", "other")
total <- "turnout"
id <- "precinct"
```

We start by using `dedupe_precincts()` to check that the data does not contain any 
duplicate precincts.

```{r}
gwinnett <- dedupe_precincts(
  data = gwinnett,
  id_cols = id
)
```
We got a warning indicating that some rows of the data do appear to belong to 
the same precinct. The function returned a column identifying problematic
rows. Let's take a look at these.

```{r}
gwinnett %>%
  filter(duplicate)
```

Here we see that precinct `035` appears to have entered the data twice. This anomaly has at least two possible explanations:

1. For most elections, precincts report their own results and data entry is often done by hand. This can lead to small mistakes like a precincts' vote tallies being split over two columns.

2. Election results databases typically contain the resutls of many elections that took place in different jurisdictions and at different times. Often when extracting the results of a particular election, one might accidentally extract some results from other elections, leading to repeated precinct IDs. 

In this case, we have confidence that this duplicate has emerged through an error in the precinct reporting. We can combine the two columns using this code:

```{r}
missing_inds <- which(gwinnett$duplicate)
columns_to_add <- c(total, cands)
gwinnett[missing_inds[1], columns_to_add] <-
  gwinnett[missing_inds[1], columns_to_add] +
  gwinnett[missing_inds[2], columns_to_add]
gwinnett <- gwinnett[-missing_inds[2]]
gwinnett[missing_ids[1], ]
```
Now we see the vote totals from the previous two columns have been summed together in this new row. We can double check that this eliminated all duplicates be rerunning `dedupe_precints()`

```{r}
gwinnett <- dedupe_precincts(
  data = gwinnett,
  id_cols = id
)
```




```{r}

```

We have this information for the forty four precincts in the county. 
